---
title: "Pubs and bars by deprivation"
author: "Jon Minton"
date: '2022-05-27'
output: html_document
---


# Aims

This document aims to see how straightforward (or otherwise) it would be to plot the locations of pubs and bars in Scotland alongside a choropleth indicating SIMD scores for different regions. 


# Packages and links

I created the following link by going to the preferred API for openstreetmap, called [overpass](https://overpass-turbo.eu/), and on the Wizard writing "pubs or bars in Scotland". 

I *think* this is a json object. 

```{r}
pub_bar_link <- "https://overpass-api.de/api/interpreter?data=%2F*%0AThis%20has%20been%20generated%20by%20the%20overpass-turbo%20wizard.%0AThe%20original%20search%20was%3A%0A%E2%80%9Cbar%20or%20pub%20in%20Scotland%E2%80%9D%0A*%2F%0A%5Bout%3Ajson%5D%5Btimeout%3A25%5D%3B%0A%2F%2F%20fetch%20area%20%E2%80%9CScotland%E2%80%9D%20to%20search%20in%0Aarea%28id%3A3600058446%29-%3E.searchArea%3B%0A%2F%2F%20gather%20results%0A%28%0A%20%20%2F%2F%20query%20part%20for%3A%20%E2%80%9Cbar%E2%80%9D%0A%20%20node%5B%22amenity%22%3D%22bar%22%5D%28area.searchArea%29%3B%0A%20%20way%5B%22amenity%22%3D%22bar%22%5D%28area.searchArea%29%3B%0A%20%20relation%5B%22amenity%22%3D%22bar%22%5D%28area.searchArea%29%3B%0A%20%20%2F%2F%20query%20part%20for%3A%20%E2%80%9Cpub%E2%80%9D%0A%20%20node%5B%22amenity%22%3D%22pub%22%5D%28area.searchArea%29%3B%0A%20%20way%5B%22amenity%22%3D%22pub%22%5D%28area.searchArea%29%3B%0A%20%20relation%5B%22amenity%22%3D%22pub%22%5D%28area.searchArea%29%3B%0A%29%3B%0A%2F%2F%20print%20results%0Aout%20body%3B%0A%3E%3B%0Aout%20skel%20qt%3B" 
```

I'm going to use the packages tidyverse, sf, tmap, and geojsonsf

```{r}
pacman::p_load(
  tidyverse, tmap, sf, #geojsonsf
  osmdata             
               )
```


# Getting the data from json to sf 

The (new) challenge will be seeing if I can convert the contents of the above link into an sf object, containing points and metadata for each of the bars and pubs 

```{r eval = FALSE}
barpub_point <- geojson_sf(pub_bar_link)

# fails: no type member at object index - - invalid GeoJSON

barpub_point <- geojson_sf(pub_bar_link, expand_geometries = TRUE)

# fails: no type member at object index - - invalid GeoJSON


# I'm instead going to see if a single object that appears to be of the right type loads as a point 

test_node <- 
  '
{
  "type": "node",
  "id": 25346762,
  "lat": 55.9349346,
  "lon": -3.1790559,
  "tags": {
    "addr:city": "Edinburgh",
    "addr:housenumber": "233-235",
    "addr:street": "Causewayside",
    "amenity": "pub",
    "name": "Old Bell Inn",
    "phone": "+44 131 668 1573",
    "real_ale": "yes",
    "source": "survey",
    "website": "https://www.oldbelledinburgh.co.uk/",
    "wheelchair": "limited"
  }
}
'

a_point <- geojson_sf(test_node)


```

This failed. The json isn't the kind of geojson that the above package is intending to work with.

Instead it appears there's a dedicated package, [osmdata](https://cran.r-project.org/web/packages/osmdata/vignettes/osmdata.html), which could/should be used instead. 

```{r}
# I'm going to start by seeing if the examples in the tutorial above work, then try adapting them to getting pubs and bars in Scotland 

x <- opq(bbox = "greater london uk") %>% 
  add_osm_feature(key = "highway", value = "motorway") %>% 
  osmdata_sf()

x

x$osm_points
plot(x$osm_points)

qtm(x$osm_points)
```

That seems to work okay, so long as I remember the sf objects are inside the main list object

Let's try to adapt the above to the query I already had 

```{r}
q1 <- opq('Scotland') %>% 
  add_osm_feature(key = 'amenity', value = 'pub')
pubs <- osmdata_sf(q1)

q2 <- opq('Scotland') %>% 
  add_osm_feature(key = "amenity", value = "bar")
bars <- osmdata_sf(q2)

pubs_or_bars <- c(pubs, bars)

```
Let's now try to plot this 

```{r}
#  tm_shape(pubs_or_bars$osm_points) + tm_dots() +# The points themselves
#  tm_shape(pubs_or_bars$osm_polygons) +tm_polygons()

# It looks like perhaps some pubs have been defined as polygons instead of points 

tm_shape(pubs_or_bars$osm_points) + tm_dots()
```
This includes all points in the bbox that contains Scotland, so bits of Northern England and Ireland/Northern Ireland as well. However hopefully we should be able to clip the points data .

Let's load in the SIMD data I downloaded earlier 


```{r}
simd <- read_sf("big_data/SG_SIMD_2020/SG_SIMD_2020.shp")

simd

```

It looks from [here](https://mattherman.info/blog/point-in-poly/) that getting points in polygons is a single line 

```{r}
#pubbars_in_scotland <- st_join(pubs_or_bars$osm_points, simd, join = st_within)

# doesn't work as the CRS are different

#st_crs(pubs_or_bars$osm_points)
#st_crs(simd)

# Will the following work? 

st_transform(pubs_or_bars$osm_points, st_crs(simd))

st_crs(pubs_or_bars$osm_points)

st_crs(pubbars_in_scotland)
st_crs(simd)
dim(pubbars_in_scotland)
dim(pubs_or_bars$osm_points)

#hmm, this doesn't seem to have reduced the number of rows...


```

```{r}
tm_shape(pubbars_in_scotland) + tm_dots()
# Nope, still includes England and Ireland(s)

# Let's try st_intersection 

pubbars_in_scotland2 <- st_intersection(pubs_or_bars$osm_points, simd)

tm_shape(pubbars_in_scotland2) + tm_dots()
# Nope! 

#check simd again

tm_shape(simd) + tm_polygons() 
# Takes a while but works 

tm_shape(simd) + tm_polygons() + 
  tm_shape(pubbars_in_scotland) + tm_dots()

# Idon't see any points

tm_shape(simd) + tm_polygons() + 
  tm_shape(pubbars_in_scotland) + tm_dots()


```
Could it be the points are presnt but just too small?! To see this let's try the interactivce mode 


```{r}

tmap_mode("view")

tm_shape(simd) + tm_polygons() + 
  tm_shape(pubbars_in_scotland) + tm_dots()


```

